Incentivizing Exploration by Heterogeneous Users

### 2018-02-07 algorithm.tex
changed some notations in the appendix and did some edits for the regret proof (not fixed)

\begin{proof}
As before, let $\LatePhase = \max (50, \frac{92.16 \sigma^4}{x^4})$.
By Lemma~\ref{lem:phase-length}, each phase $s = 1, \ldots, \LatePhase$
in expectation lasts at most $\ARMNUM s$ steps,
so the total expected regret incurred in these phases is at most
$\sum_{s=1}^{\LatePhase} \ARMNUM \MAXR s = O(\ARMNUM \MAXR \LatePhase^2)$.

Regret incurred after the first \LatePhase phases can arise in two ways:
(1) an agent was incentivized to pull a suboptimal arm, or
(2) an agent myopically pulled a suboptimal arm.
By Lemma~\ref{lem:numP}, Algorithm~\ref{alg:basic-incentivizing}
incentivizes agents at most 
$O \left( \frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)$
times in expectation, each time causing regret at most \MAXR,
for a total expected regret of at most
$O \left( \frac{\MAXR \ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)$.
For the rest of the proof, we focus on the regret incurred when agents
pull arms myopically and make mistakes.

We distinguish between agents incurring large regret
(which requires severe misestimates of arm locations;
such misestimated are exponentially unlikely to occur), 
and agents incurring small positive regret,
which requires these agents to be almost tied in their preference for
the best arm.
To be more precise, we will define (with hindsight) a time-dependent
cutoff $g(t)$, and consider a regret exceeding $g(t)$ large.

We first consider agents \AgV incurring positive regret less than $g(t)$.
The second-best arm \Second{\AgV} of \AgV must satisfy
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq g(t)$.
By Lemma~\ref{lem:sdelta}, the total measure of such agents
is $\AlmostTied{g(t)} \leq \TieDensity \cdot g(t)$,
and each incurs regret at most $g(t)$.
Summing over all time steps, the total regret from such agents is at most
$\sum_{t = 1}^T \TieDensity \cdot g^2(t)$.

To bound the regret incurred in time steps with large regret,
let the random variable $t_s$ be the final time step in phase $s$,
for each $s$.
\dkcomment{Is $t_s$ better or $T_s$?}
Define the stopping times $\tau_{s}^{k}$ to be the \Kth{k}
time step in the \Kth{s} phase,
with $\tau_{s}^{k} = \infty$ when phase $s$ has fewer than $k$ steps,
i.e., $k > t_{s}-t_{s-1}$.
For $t = \tau_{s}^{k}$, by Corollary~\ref{cor:right-choice}
(applied with $x = \frac{g^2(\tau_s^k)}{4 \Diam^2 d^2}$),
the probability of incurring regret exceeding $g(t)$ is at most
$24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s g(\tau_{s}^{k})^2}{64 \Diam^2 d^2\sigma^2}\right)$.
Because each arm must have been pulled at least $s$ times by the time
the algorithm reaches phase $s$,
\dkcomment{Really only $s-1$; we have to be careful with those $\pm 1$.}
we can bound $\tau_{s}^{k} \geq s$.
\dkcomment{We could bound it by $\ARMNUM s$, if that helps get
  slightly better bounds.}
Because the regret never exceeds \MAXR,
the expected (large) regret for the \Kth{k} time step of phase $s$
is bounded above by
$24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s g(s)^2}{64 \Diam^2 d^2\sigma^2}\right) \cdot \MAXR$.
Summing over all time steps $t$, we get that the total regret from
large-regret steps is at most

\begin{align*}
\lefteqn{\sum_{s = \LatePhase}^T \sum_{t=t_s}^{t_{s+1}-1}
  24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s g^2(s)}{64 \Diam^2 d^2\sigma^2}\right) \cdot \MAXR}
\\ & =
\sum_{s = \LatePhase}^T 
  24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s g^2(s)}{64 \Diam^2 d^2\sigma^2}\right) \cdot \MAXR
  \cdot \Expect{\mbox{Length of phase } s}
\\ & \leq
\sum_{s = \LatePhase}^T 
  24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s g^2(s)}{64 \Diam^2 d^2\sigma^2}\right) \cdot \MAXR
  \cdot \ARMNUM s,
\end{align*}
where the inequality was an application of Lemma~\ref{lem:phase-length}.
Adding all four terms, and starting the sum at 1 instead of \LatePhase,
the cumulative regret up to time $T$ is at most

\begin{multline*}
O(\ARMNUM \MAXR \LatePhase^2) + 
O \left( \frac{\MAXR \ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right) +
\\
+ \sum_{s = 1}^T 
  24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s g^2(s)}{64 \Diam^2 d^2\sigma^2}\right) \cdot \MAXR
  \cdot \ARMNUM s,
+ \sum_{t=1}^{T}\TieDensity \cdot g^2(t).
\end{multline*}

The first line is independent of $T$, and we now choose
$g^2(t) = \frac{2\log(T) \cdot 64 \Diam^2 d^2 \sigma^2}{1.8 t}$
with the goal of minimizing the second line.
With this choice, the second line becomes

\begin{align*}
\lefteqn{\sum_{s=1}^{T} 24\ARMNUM^2 d s \MAXR \exp\left(-2\log(T)\right)  +
          \frac{128\Diam^2 d^2\sigma^2
          \TieDensity\log(T)}{1.8}\sum_{t=1}^{T}\frac{1}{t}}
\\ & \leq
24\ARMNUM^2 d R\frac{T(T-1)}{2T^2}  + 71.12 \Diam^2 d^2\sigma^2
          \TieDensity\log(T)(\log(T)+1)
  \\ & \leq
12 \ARMNUM^2 d R  + 71.12 \Diam^2 d^2\sigma^2 \TieDensity\log(T)(\log(T)+1). 
\end{align*}

>>>>

\begin{proof}
Regret can arise in two ways:
(1) an agent was incentivized to pull a suboptimal arm, or
(2) an agent myopically pulled a suboptimal arm.
By Lemma~\ref{lem:numP}, Algorithm~\ref{alg:basic-incentivizing}
incentivizes agents at most 
$O \left( \frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)$
times in expectation, each time causing regret at most \MAXR,
for a total expected regret of at most
$O \left( \frac{\MAXR \ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)$.
For the rest of the proof, we focus on the regret incurred when agents
pull arms myopically and make mistakes.

We distinguish between agents incurring large regret
(which requires severe misestimates of arm locations;
such misestimated are exponentially unlikely to occur), 
and agents incurring small positive regret,
which requires these agents to be almost tied in their preference for
the best arm.
To be more precise, we will define (with hindsight) a phase-dependent
cutoff $g(s)$, and consider a regret exceeding $g(s)$ large.

We first consider agents \AgV incurring positive regret less than $g(s)$.
The second-best arm \Second{\AgV} of \AgV must satisfy
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq g(s)$.
By Lemma~\ref{lem:sdelta}, the total measure of such agents
is $\AlmostTied{g(s)} \leq \TieDensity \cdot g(s)$,
and each incurs regret at most $g(s)$.
Summing over all time steps, the expected total regret from such agents is at most
\begin{align}
&E[\sum_{s=1}^{T}\sum_{t=t_{s-1}+1}^{t_{s}} \TieDensity \cdot g^2(s)] \leq \sum_{s=1}^{T}\TieDensity\cdot g^2(s)Ns. \nonumber 
\end{align}

To bound the regret incurred in time steps with large regret,
let the random variable $t_s$ be the final time step in phase $s$,
for each $s$.
\dkcomment{Is $t_s$ better or $T_s$?}
Define the stopping times $\tau_{s}^{k}$ to be the \Kth{k}
time step in the \Kth{s} phase,
with $\tau_{s}^{k} = \infty$ when phase $s$ has fewer than $k$ steps,
i.e., $k > t_{s}-t_{s-1}$.
For $t = \tau_{s}^{k}$, we know
\begin{align}
&\Prob{\text{incur regret exceeding g(s)}} \nonumber \\
\leq & \Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}}.\nonumber 
\end{align}

Because the regret never exceeds \MAXR,
the expected (large) regret for the \Kth{k} time step of phase $s$
is bounded above by
$\Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \cdot \MAXR$.
Summing over all time steps $t$, we get that the total regret from
large-regret steps is at most

\begin{align*}
\lefteqn{\sum_{s = 1}^T \sum_{t=t_s}^{t_{s+1}-1}
  \Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \cdot \MAXR}
\\ & =
\sum_{s = 1}^T 
  \Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \cdot \MAXR
  \cdot \Expect{\mbox{Length of phase } s}
\\ & \leq
\sum_{s = 1}^T 
 \Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \cdot \MAXR
  \cdot \ARMNUM s,
\end{align*}
where the inequality was an application of Lemma~\ref{lem:phase-length}.
Adding all three terms,
the cumulative regret up to time $T$ is at most

\begin{align*}
&O \left( \frac{\MAXR \ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right)
\\
+ & \sum_{s = 1}^T 
  \Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \cdot \MAXR
  \cdot \ARMNUM s,
+ \sum_{s=1}^{T}\TieDensity \cdot g^2(s)\cdot Ns.
\end{align*}

The first line is independent of $T$, and we now choose
$g^2(s) = \frac{288\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$
with the goal of minimizing the second line.
With this choice and same definition of $S_{s}$ as in the proof of Lemma~\ref{lem:round-prob}, 
\begin{align*}
&\Prob{\exists i,j, s.t. |\ArmE{t}{i}{j} - \Arm{i}{j}|\geq \frac{g(s)}{2\Diam d}} \\
\leq &\sum_{i,j}\Prob{|S_s|\geq 3\sqrt{\frac{2}{1.8}s\log(T)}} \\
\leq &\sum_{i,j}\Prob{|S_s|\geq \sqrt{0.6s\log(\log_{1.1}(s)+1)+\frac{2}{1.8}\log(T)s}} \leq 24\ARMNUM d \cdot \exp(-2\log(T)),
\end{align*}
where the second to last inequality is because the following technical inequality, whose proof we include in the appendix.

\begin{lemma} \label{lem:n1-inequality}
For $2\leq s\leq T$, we have $3\sqrt{\frac{2}{1.8}s\log(T)} \geq \sqrt{0.6s\log(\log_{1.1}(s)+1)+\frac{2}{1.8}\log(T)s}$.
\end{lemma}

\bccomment{I believe the calculation up to this line should be correct now. The final equation we want to minimize is $\sum_{s=1}^{T}[\exp(-s*g^2(s))+g^2(s)*s]$, what should be our choice of $g(s)$?}

Therefore, the second line becomes

\begin{align*}
\lefteqn{\sum_{s=1}^{T} 24\ARMNUM^2 d s \MAXR \exp\left(-2\log(T)\right)  +
          \frac{128\Diam^2 d^2\sigma^2
          \TieDensity\log(T)}{1.8}\sum_{s=1}^{T}\frac{1}{s}}
\\ & \leq
24\ARMNUM^2 d R\frac{T(T-1)}{2T^2}  + 71.12 \Diam^2 d^2\sigma^2
          \TieDensity\log(T)(\log(T)+1)
  \\ & \leq
12 \ARMNUM^2 d R  + 71.12 \Diam^2 d^2\sigma^2 \TieDensity\log(T)(\log(T)+1). 
\end{align*}
